Predicting Tags for StackOverflow Posts

### 2013-02-10 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 9b20364a4ea81ec508218b529fe3b4654342e0b4 [formerly 8b23a2e0941e6297285b818c2b5e4dca03ced9b4]
Former-commit-id: 549049bea4a6e08c598cf7221559a1917bcc77c6

The base R glm function (with 'binomial' argument) was used to perform the logistic regression.
The parameters were optimized to best predict categorization (chosen tag, not chosen tag) from the four predictor variables:
prior tag occurrence, word co-occurrence activation for title and body words, and activation offset for the post.
Further discussion and operationalization of the predictor variables is included in the Model section.

@@

Here, the $_{i}$ subscript denotes activation for a particular tag i, and the $_{j}$ subscript is for context.
$A_{i}$ is total activation for tag i.
Note the similarities between the StackOverflow model and ACT-R's retrieval equations (base level and contextual component, log odds of prior being adjusted by association strength with context).

@@

Base-level activation ($B_{i}$) is the log prior odds of observing tag i in the world.
So the tag with the highest frequency (C\#) would have the highest base-level activation.
Tag frequency drops off sharply for this dataset.
The top 1\% of tags based on frequency of use account for approximately 60\% of total tag instances.
This is why accounting for tag frequency (by including ACT-R's base-level activation component) is so important for the StackOverflow dataset.
Otherwise without context, the likelihood of choosing C\# as a tag would be the same as choosing ora-00997 (one of the many tags with only a single occurrence of use).
One might argue that with context, it's unlikely that the words associated with ora-00997 will occur again, so pure context would be able to still often predict C\#.
However, the ACT-R theory asserts that pure context is often not enough to overcome high base-level activations (e.g., it can be difficult to break a bad habit).

@@

Contextual activation can be thought of as the amount of odds adjustment to base levels, given the current context $_{j}$.
Contextual activation has two components: strength of association ($S_{ji}$) and attentional weight ($W_{j}$).
Strength of association measures the relatedness of a contextual element $_{i}$ (e.g., the word PHP in the title of a post) with a tag $_{j}$ (e.g., the tag PHP5).

@@

The second component of contextual activation, attentional weighting ($W_{j}$), can be thought of as the amount of attentional resources that are dedicated to contextual element $_{j}$.
Total attentional weighting ($W$) is bound, most often set to 1, and reflects the fact that our attentional resources are limited.
For the StackOverflow model, the attentional weighting was left unbound, and allowed to be calibrated by the logistic regression model.
This was done for two main reasons.
First, the researcher was curious to see what value the logistic regression model would give for $W$.
Second, a non-ACT-R standard measure of attentional weighting was used ($E_{j}$), so it seemed appropriate to allow $W$ to vary alongside exploring this new weighting measure.

>>>>

The base R glm function (with 'binomial' argument) was used to perform the logistic regression.
The parameters were optimized to best predict categorization (chosen tag, not chosen tag) from the four predictor variables:
The tag's prior occurrence frequency, word co-occurrence activation for title and body words, and activation offset for the post.
Further discussion and operationalization of the predictor variables is included in the Model section.

@@

Here, the $i$ subscript denotes activation for a particular tag $i$, and the $j$ subscript is for context.
$A_{i}$ is total activation for tag $i$.
Note the similarities between the StackOverflow model and ACT-R's retrieval equations (base level and contextual component, log odds of prior being adjusted by association strength with context).

@@

Base-level activation ($B_{i}$) is the log prior odds of observing tag $i$ in the world.
So the tag with the highest frequency (C\#) would have the highest base-level activation.
Tag frequency drops off sharply for this dataset.
The top 1\% of tags based on frequency of use account for approximately 60\% of total tag instances.
This is why accounting for tag frequency (by including ACT-R's base-level activation component) is so important for the StackOverflow dataset.
Otherwise without context, the likelihood of choosing C\# as a tag would be the same as choosing ora-00997 (one of the many tags with only a single occurrence of use).
One might argue that with context, it's unlikely that the words associated with ora-00997 will occur again, so pure context would be able to still often predict C\#.
However, the ACT-R theory asserts that pure context is often not enough to overcome high base-level activations (e.g., it can be difficult to break a bad habit).

@@

Contextual activation can be thought of as the amount of odds adjustment to base levels, given the current context $j$.
Contextual activation has two components: strength of association ($S_{ji}$) and attentional weight ($W_{j}$).
Strength of association measures the relatedness of a contextual element $i$ (e.g., the word PHP in the title of a post) with a tag $j$ (e.g., the tag PHP5).

@@

The second component of contextual activation, attentional weighting ($W_{j}$), can be thought of as the amount of attentional resources that are dedicated to contextual element $j$.
Total attentional weighting ($W$) is bound, most often set to 1, and reflects the fact that our attentional resources are limited.
For the StackOverflow model, the attentional weighting was left unbound, and allowed to be calibrated by the logistic regression model.
This was done for two main reasons.
First, the researcher was curious to see what value the logistic regression model would give for $W$.
Second, a non-ACT-R standard measure of attentional weighting was used ($E_{j}$), so it seemed appropriate to allow $W$ to vary alongside exploring this new weighting measure.
